\begin{center} \bf \sffamily \LARGE Prisoner's dilemma with stochastic payoffs\\[-1cm] \end{center}

### 2023-02-23 paper/main.tex
milestone in writing

Similar to all other theoretical models, finite stochastic processes rest on a
set of assumptions. Namely, in the game phase it is assumed that players use
strategies with finite memory~\cite{Nowak1992tit, Baek2016}. For example, in a
repeated game between two individuals the actions of the players in each turn
are often determined by the action of the co-players in the previous turn. This
assumption is common as it allows for an explicit calculation of the players'
payoffs~\cite{sigmund2010calculus}. In addition, it is
assumed that individuals play many times and with all other players before
reproduction takes. So that in the updating phase the updating payoffs, and
subsequently the fitness, of individuals us given by the average payoffs of an
infinitely repeated game against the mean distribution of types in the
population.

@@

We first consider two extreme scenarios, the classical scenario where individual
has a perfect memory of all their interactions and the alternative scenario of
limited memory where individuals update their strategies only based on the very
last payoff they obtained. We demonstrate the effects of limited updating payoff
memory using a pairwise comparison process and the well studied game of the
prisoner's dilemma. We observe that individuals with limited memory tend to
adopt less generous strategies and they achieve less cooperation when
interacting in a prisoner's dilemma. We obtain similar results when we consider
that individuals update their strategies based on more information. More
specifically, up to the last two payoffs they obtained when interacting with up
to two different members of the population.

@@

A pairwise comparison process starts with assigning all
individuals of the population the same strategy. Each elementary time step of
the process consists of the mutation phase, the game phase and the update phase.
In the game phase individuals are matched in pairs and that they participate in
a repeated 2 persons donation game; a special case of the prisoner's dilemma. In
the donation game there are two actions: cooperation (\(C\)) and defection
(\(D\)). By cooperating a player provides a benefit \(b\) to the other player at
their cost \(c\), with \(0 < c < b\). Thus the payoffs for a player in each turn
are,

@@

In repeated games there are infinite many strategies, however, similar to the
literature we will assume that individuals use reactive strategies. A reactive strategy considers only the
previous action of the other player, and thus, a reactive strategy \(s\) can be
written as a three-dimensional vector \(s=(y, p, q)\). The parameter \(y\) is
the probability that the strategy opens with a cooperation and \(p\), \(q\) are
the probabilities that the strategy cooperates given that the opponent
cooperated and defected equivalently. The play between a pair of reactive
strategies can be model as a Markov process with the transition matrix \(M\),

@@

and the stationary vector \(\mathbf{v}(s_1,s_2)\)
which is the solution to \(\mathbf{v}(s_1,s_2) \cdot M 
= \mathbf{v}(s_1,s_2)\).

@@

We will mainly assume that mutations are rare (\(\mu \rightarrow 0\)). In fact,
so rare that only two different strategies can be present in the population at
any given time. However, we also show in section~\ref{section:mutation} that the
main results hold for \(\mu \neq 0\). The case of low mutation is vastly adopted because it allows us to explicitly
calculate the fixation probability of a newly introduced mutant. At each step
one individual adopts a mutant strategy randomly selected from the set of
feasible strategies. The fixation probability \(\phi_{M}\) of the mutant
strategy can be calculated explicitly,

\begin{equation}\label{eq:appendix_fixation_probability}
    \varphi = \frac{1}{1+\sum\limits_{i=1}^{N-1}\prod\limits_k^i \frac{\lambda^-_k}{\lambda^+_k}},
\end{equation}

where \(\lambda^-_k, \lambda^+_k\) are the probabilities that the number of
mutants decreases and increases respectively, and \(k\) is the number of mutants. Depending on the fixation
probability \(\phi_{M}\) the mutant either fixes (becomes the new resident) or
goes extinct. Regardless, in the elementary time step another mutant strategy is
introduced to the population. We iterate this elementary population updating
process for a large number of mutant strategies and we record the resident
strategies at each time step.

The probabilities \(\lambda^-_k \text{ and } \lambda^+_k\) depend on the
updating payoffs of the mutant and the resident strategies. In the next
section we discuss the fitness of individuals in the case of perfect and
limited memory.

@@

In the perfect memory case an individual updates based on the average payoff
against each other member of the population, otherwise as expected payoffs. In
an infinitely repeated game ($\delta \rightarrow 1$) the payoff of a reactive
strategy \(s_1\) against the reactive strategy \(s_2\) can explicitly be
calculated using the stationary vector \(\mathbf{v}\) and the payoffs' vector as,

\[\langle\mathbf{v}(s_1,s_2),\mathbf{u}\rangle \quad.\]

In a population of size \(N\), there are \(k\) mutants and \(N - k\) residents,
whose strategies we denote respectively as \(s_M =(y_M, p_M, q_M)\) and \(s_R =
(y_R, p_R, q_R)\). The expected payoffs of a resident (\(\pi_R\)) and for a
mutant (\(\pi_M\)) are give by,

@@

\begin{align}
  \lambda^-_k \!=\!\rho(\pi_M, \pi_R) \quad \text{ and } \quad \lambda^+_k \!=\!\rho(\pi_R, \pi_M).
\end{align}

@@

In this case of limited memory we initially define the probability that a
reactive strategy receives the payoff $u\!\in\! \mathcal{U}$ in the very last
round of the game. This is given by Proposition~\ref{proposition:last_round}.
The proof can be found in the Appendix [ref].

@@

We consider the case where both the role model and the learner estimate their
fitness after interacting with a single member of the population. There are five
possible pairings at each step. They interact with other with a probability
\(\frac{1}{N - 1}\), and thus they do not interact with other with a probability
\(1 - \frac{1}{N - 1}\). In the latter case, each of them can interact with
either a mutant or a resident. Both of them interact with a mutant with a
probability $\frac{(k-1)(k-2)}{(N-2)(N-3)}$ and both interact with a resident
with a probability $\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$. The last two possible
pairings are that either of them interacts with a resident whilst the other
interacts with a mutant, and this happens with a probability
$\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$. The probability that the respective last
round payoffs of two players \(s_1, s_2\) are given by $u_1$ and $u_2$ as,

@@

The first term on the right side corresponds to the case that the learner and
the role model happened to be matched during the game stage, which happens with
probability $\frac{1}{(N\!-\!1)}$. In that case, we note that only those payoff
pairs can occur that are feasible in a direct interaction,
$\splitatcommas{(u_1,u_2)\in \mathcal{U}^2_F:=\big\{ (r, r), (s, t), (t, s), (p, p)
\big\}}$, as represented by the respective indicator function. Otherwise, if the
learner and the role model did not interact directly, we need to distinguish
four different cases, depending on whether the learner was matched with a
resident or a mutant, and depending on whether the role model was matched with a
resident or a mutant.

The probability that the number of mutants increases, and decreases respectively,
by one is now given by,

\begin{align}
\lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M})\cdot \rho(u_{R},u_{M}), \\
\lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M})\cdot \rho(u_{M},u_{R}).
\end{align}

@@

% We use
% $x(u_R,u_M)$ to denote the probability that the randomly chosen resident
% obtained a payoff of $u_R$ in the last round of his respective game, and that
% the mutant obtained a payoff of $u_M$.

@@

In order to account for the effect of the updating payoffs we simulate the
evolutionary process and record which strategies the players adopt over time
based on perfect memory payoffs and limited memory payoffs. We run two
independent runs for each approach and we vary the value of benefit \(b\).
Figure~\ref{fig:expected_and_stochastic_for_donation} shows the simulation
results. It depicts the evolving conditional cooperation probabilities $p$ and
$q$. The discount factor~$\delta$ is comparably high, thus we do not report the
opening move \(y\) as it is a transient effect. The upper panel correspond to
the standard scenario considered in the literature, it considers players who use
expected payoffs to update their strategies. The bottom panel shows the scenario
considered herein, in which players update their strategies based on their last
round's payoff.

The figure suggests that when updating is based on perfect memory payoffs
players tend to be more generous and more cooperative. The $q$-values of the
resident strategies are on average higher in the case of perfect memory. Thus,
players will occasionally forgive a defection more often if their fitness
depends on interacting with every member of the population, and this effects
becomes more abundant as the value of the benefit increases. In comparison, in
the case of limited memory the evolved generosity is independent of the values
of benefit, and they are less forgiving. We also calculate the average
cooperation rate for each simulation which is the average cooperation rate
within the resident population. In the case of the perfect memory payoffs the
average cooperation rate is strictly higher than that of the last round payoffs.
The difference based on the two methods is statistically significant, and in the
case of $b=10$ the average cooperation of resident strategies drops from 97\% to
57\%.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=\textwidth]{static/donation_expected_last_round_summary_results.pdf}
    \caption{{\bf Evolutionary dynamics under perfect and limited payoff memory.}
    ({\bf Schematic illustrations})
    On the left panels we show schematic illustrations of the perfect memory and
    the limited memory cases. The shaded background denotes the game phase
    information that an individual considered when updating strategies. In the
    case of perfect memory the entire region is shaded and in the case of
    limited memory only one turn with a single member of the population.
    ({\bf Simulations})
    We have run four simulations of the pairwise comparison process
    for $t\!=\!10^7$ time steps. For each time step,
    we have recorded the current resident population ($y,p,q$). Since
    simulations are run for a relatively high continuation probability of
    $\delta\!=\!0.999$, we do not report the players' initial cooperation
    probability $y$. The graphs show how often the resident population chooses
    each combination ($p,q$) of conditional cooperation probabilities in the
    subsequent rounds. We also report the evolved cooperation rate which is
    calculated as the average cooperation rate within the resident population.
    ({\bf Perfect Memory}) In the case of low benefit
    the resident population either consists of defectors (with $p\!\approx\!q\!\approx\!0$) or of
    conditional cooperators. Conditional cooperators, or otherwise known as generous
    tit for tat, are a set of strategies that always cooperate following a cooperation
    ($p\!\approx\!1\!$) and cooperate with a probability
    $q$ given that the co-player has defected. $q$ denotes the generosity
    of a player. The resident population applies a conditional cooperator strategy for which
    $q\!\le\!1\!-\!c/b\!=\!0.67$. This is true also for the case of high benefit.
    In this case the population mainly consists of conditional
    cooperators of the form ($p\!\approx\!1\!, q\!\le\!1\!-\!1/10\!=\!0.9$).
    In the Appendix we show that a conditional cooperator needs to be
    of the form ($p\!\approx\!1\!, q\!\le\!1\!-\!c/b$) to not be invaded by
    defecting strategies. A higher generosity in the population results to a
    higher average cooperation rate. The average cooperation rate
    increases from a 52\% for \(b\) to 98\% for \(b=10\).
    ({\bf Limited Memory}) When players
    update their strategies based on their realized payoffs in the last round,
    there are two different predominant behaviors regardless of the benefit value. The resident population
    either consists of defectors (with $p\!\approx\!q\!\approx\!0$) or of
    conditional cooperators. The maximum level of $q$
    consistent with stable cooperation is somewhat smaller compared to the
    perfect memory setting, $q\!<\!0.5$. Namely, in the Appendix we show
    that regardless the value of benefit, in the case of limited payoff memory,
    an individual needs generosity smaller of $0.5$
    to repel defectors. The plots
    are fairly similar regardless of the benefit, and the evolved cooperation
    rate only sightly increases from 37\% to 51\%. Parameters: $N\!=\!100$, $c\!=\!1$, $\beta\!=\!1$,
    $\delta\!=\!0.999$.}
    \label{fig:expected_and_stochastic_for_donation}
\end{figure}

We further explore the effect that the value of benefit \(b\) and the strength
of selection \(\beta\) have on the generosity \(q\) and to the evolved
cooperation rate, Figure~\ref{fig:cooperation_rate_over_benefit_and_beta}. The
figure suggests that expected payoffs always yield a higher cooperation rate. In
the case of expected payoffs we observe that the cooperation rate increases as
the value of the benefit gets higher. In comparison for the limited memory
payoffs, the cooperation rate remains unchanged at approximately 50\% once
\(b=5\). We also investigate the effect of the strength of selection $\beta$.
For weak selection, \(\beta < 1\), we observe that the two methods yield similar
results, however, as \(\beta\) increases there is variation in the evolving
populations. In the case of expected payoffs the resident populations become
more cooperative as \(\beta\) increases, whereas in the case of limited memory
payoffs, the resident populations become more defective.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=\textwidth]{static/cooperation_rate_over_b_and_beta.pdf}
    \caption{{\bf The evolution of cooperation and generosity for different values
    of benefit (A) and strength of selection (B).} We report the
    average cooperation and the average reciprocity. The average cooperation rate
    is the average cooperation rate
    within the resident population. For the average reciprocity we check for the
    residents that ($p \approx 1$) and we take the average of theirs $q$s ({\bf A})
    We vary the benefit of defection $b$, and we report both the average
    cooperation and average generosity. In all cases, expected payoffs appear to
    overestimate the average cooperation rate the population achieves. As
    expected, since the average generosity over the different values of $beta$
    remains the same ($q \approx 0.5$), and a result so does the average
    cooperation. ({\bf B}) We vary the selection strength $\beta$. In all
    cases, stochastic payoff evaluation tends to reduce the evolving cooperation
    rates. Unless explicitly varied, the parameters of the simulation are
    $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$.
    Simulations are run for $t\!=\!5\times 10^7$ time steps for each parameter
    combination.}\label{fig:cooperation_rate_over_benefit_and_beta}
\end{figure}

@@

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\textwidth]{static/more_memory_summary_results.pdf}
  \caption{{\bf Average cooperation rates for different updating payoffs.} (From
  right to left) updating Payoffs based  on perfect memory, limited memory,
  the last round payoff of two interactions,
  the last two rounds payoff of one interaction and 
  the last two round payoffs of two interactions.({\bf High Benefit})
  ({\bf Low Benefit}) We vary the selection strength $\beta$. In all
  cases, stochastic payoff evaluation tends to reduce the evolving cooperation
  rates. Unless explicitly varied, the parameters of the simulation are
  $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$.
  Simulations are run for $t\!=\!5\times 10^7$ time steps for each parameter
  combination.}\label{fig:cooperation_rate_updating_payoffs}
\end{figure}

>>>>

Similar to all other theoretical models, finite stochastic processes rest on a
set of assumptions. Namely, in the game phase it is assumed that players use
strategies with finite memory~\cite{Nowak1992tit, Baek2016}. For example, in a
repeated game between two individuals the actions of the players in each turn
are often determined by the action of the co-players in the previous turn. This
assumption is common as it allows for an explicit calculation of the players'
payoffs~\cite{sigmund2010calculus}. In addition, it is
assumed that individuals play many times and with all other players before
reproduction takes place. So that in the updating phase the updating payoffs, and
subsequently the fitness, of individuals is given by the average payoffs of an
infinitely repeated game against the mean distribution of types in the
population.

@@

We first consider two extreme scenarios, the classical scenario where an individual
has a perfect memory of all their interactions and the alternative scenario of
limited memory where individuals update their strategies only based on the very
last payoff they obtained. We demonstrate the effects of limited updating payoff
memory using a pairwise comparison process and the well studied game of the
prisoner's dilemma. We observe that individuals with limited memory tend to
adopt less generous strategies and they achieve less cooperation when
interacting in a prisoner's dilemma. We obtain similar results when we consider
that individuals update their strategies based on more information. More
specifically, up to the last two payoffs they obtained when interacting with up
to two different members of the population.

@@

A pairwise comparison process starts with assigning all
individuals of the population the same strategy. Each elementary time step of
the process consists of the mutation phase, the game phase and the update phase.
In the game phase individuals are matched in pairs and that they participate in
a repeated 2 person donation game; a special case of the prisoner's dilemma. In
the donation game there are two actions: cooperation (\(C\)) and defection
(\(D\)). By cooperating a player provides a benefit \(b\) to the other player at
their cost \(c\), with \(0 < c < b\). Thus the payoffs for a player in each turn
are,

@@

In repeated games there are infinite many strategies, however, similar to the
literature we will assume that individuals use reactive strategies. A reactive
strategy considers only the previous action of the other player, and thus, a
reactive strategy \(s\) can be written as a three-dimensional vector \(s=(y, p,
q)\). The parameter \(y\) is the probability that the strategy opens with a
cooperation and \(p\), \(q\) are the probabilities that the strategy cooperates
given that the opponent cooperated and defected equivalently. The play between a
pair of reactive strategies ($s_1 = (y_1, p_1, q_1)$, $s_2 = (y_2, p_2, q_2)$)
can be model as a Markov process with the transition matrix \(M\),

@@

and the stationary vector \(\mathbf{v}(s_1,s_2)\)
which is the solution to \(\mathbf{v}(s_1,s_2) \times M 
= \mathbf{v}(s_1,s_2)\).

@@

For the results presented here we assume that mutations are rare (\(\mu
\rightarrow 0\)). In fact, so rare that only two different strategies can be
present in the population at any given time. However, in the Supplementary Information
Section 4 we show in that the main result hold for \(\mu \neq 0\).
The case of low mutation is vastly adopted because it allows us to explicitly
calculate the fixation probability of a newly introduced mutant. More specifically,
at each step
one individual adopts a mutant strategy randomly selected from the set of
feasible strategies. The fixation probability \(\phi_{M}\) of the mutant
strategy can be calculated explicitly,

\begin{equation}\label{eq:appendix_fixation_probability}
    \varphi_{M} = \frac{1}{1+\sum\limits_{i=1}^{N-1}\prod\limits_k^i \frac{\lambda^-_k}{\lambda^+_k}},
\end{equation}

where \(\lambda^-_k, \lambda^+_k\) are the probabilities that the number of
mutants decreases and increases respectively, and \(k\) is the number of
mutants. Depending on the fixation probability \(\phi_{M}\) the mutant either
fixes (becomes the new resident) or goes extinct. Regardless, in the elementary
time step another mutant strategy is introduced to the population. We iterate
this elementary population updating process for a large number of mutant
strategies and we record the resident strategies at each time step. The
probabilities \(\lambda^-_k \text{ and } \lambda^+_k\) depend on the updating
payoffs of the mutant and the resident strategies. In the next section we
present how they are calculated in the cases of perfect and limited memory.

@@

In the perfect memory case an individual updates based on the average payoff
against each other member of the population, otherwise as expected payoffs. In
an infinitely repeated game ($\delta \rightarrow 1$) the payoff of a reactive
strategy \(s_1\) against the reactive strategy \(s_2\) can explicitly be
calculated using the stationary vector \(\mathbf{v}\) and the payoffs' vector as
\(\langle\mathbf{v}(s_1,s_2),\mathbf{u}\rangle\). In a population of size \(N\)
there are \(k\) mutants and \(N - k\) residents. We denote the strategies of
a mutant and of a resident as \(s_M =(y_M, p_M, q_M)\) and \(s_R =
(y_R, p_R, q_R)\). The expected payoffs of a resident (\(\pi_R\)) and of a
mutant (\(\pi_M\)) are given by,

@@

\begin{align}\label{eq:perfect_memory_lambdas}
  \lambda^-_k \!=\!\rho(\pi_M, \pi_R) \quad \text{ and } \quad \lambda^+_k \!=\!\rho(\pi_R, \pi_M).
\end{align}

@@

In this case of limited memory we initially define the probability that a
reactive strategy receives the payoff $u\!\in\! \mathcal{U}$ in the very last
round of the game. This is given by Proposition~\ref{proposition:last_round}
(see Supplementary Information Section 2.2.1 for proof).

@@

In the case of limited payoffs memory both the role model and the learner
estimate their fitness after interacting with a single member of the population.
At each time step there are five possible pairings. They interact with each
other with a probability \(\frac{1}{N - 1}\), and thus they do not interact with
other with a probability \(1 - \frac{1}{N - 1}\). In the latter case, each of
them can interact with either a mutant or a resident. Both of them interact with
a mutant with a probability $\frac{(k-1)(k-2)}{(N-2)(N-3)}$ and both interact
with a resident with a probability $\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$. The last
two possible pairings are that either of them interacts with a resident whilst
the other interacts with a mutant, and this happens with a probability
$\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$. Thus we the probability that the randomly
chosen resident obtained a payoff of $u_R$ in the last round of his respective
game, and that the mutant obtained a payoff of $u_M$ as $x(u_R,u_M)$.

@@

The first term on the right side corresponds to the case that the learner and
the role model happened to be matched during the game stage, which happens with
probability $\frac{1}{(N\!-\!1)}$. The probability that the number of mutants
increases and decreases by one in the case of limited memory are now given by,

\begin{align}\label{eq:limited_memory_lambdas}
\lambda^+_k=\frac{N\!-\!k}{N} \frac{k}{N} \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M}) \rho(u_{R},u_{M}) \quad \text{and} \quad
\lambda^-_k=\frac{N\!-\!k}{N} \frac{k}{N} \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M}) \rho(u_{M},u_{R}).
\end{align}

@@

@@

In order to account for the effect of the updating payoffs we simulate the
evolutionary process and record which strategies the players adopt over time
based on perfect memory payoffs and limited memory payoffs. In the case of
perfect memory payoffs we use Eq~\ref{eq:perfect_memory_lambdas} when we
estimate the probability that a mutant fixes, and
Eq~\ref{eq:limited_memory_lambdas} in the case on limited memory. We run two
independent runs for each approach and we vary the value of benefit \(b\).
Figure~\ref{fig:expected_and_stochastic_for_donation} shows the simulation
results. It depicts the evolving conditional cooperation probabilities $p$ and
$q$. The discount factor~$\delta$ is comparably high, thus we do not report the
opening move \(y\) as it is a transient effect. The upper panel corresponds to
the standard scenario considered in the literature, it considers players who use
expected payoffs to update their strategies. The bottom panel shows the scenario
considered herein, in which players update their strategies based on their last
round's payoff.

The figure suggests that when updating is based on perfect memory payoffs
players tend to be more generous and more cooperative. The $q$-values of the
resident strategies are on average higher in the case of perfect memory. Thus,
players will occasionally forgive a defection more often if their fitness
depends on interacting with every member of the population, and this effect
becomes more notable as the value of the benefit increases. In the case of
limited memory the resident strategies are less forgiving and their forgiveness
is independent of the values of benefit. In the Supplementary Information
Sections 2.1 and 2.2.1 we show that $q$-values of a resident strategy have to be less
than \(1 - c/b\) in the perfect memory case and less than \(0.5\) in the limited
memory case. The generosity for the perfect memory case is always higher for a
benefit value higher than 2, and it increases as the benefit value increases.
This leads to a more cooperative population. We calculate the average
cooperation rate for each simulation which is the average cooperation rate
within the resident population. In the case of the perfect memory payoffs the
average cooperation rate is strictly higher than that of the last round payoffs,
and the difference is most for $b=10$. The average cooperation of resident
strategies drops from 97\% to 57\%. This indicates that the expected payoffs
overestimate the evolved cooperation.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=\textwidth]{static/donation_expected_last_round_summary_results.pdf}
    \caption{{\bf Evolutionary dynamics under perfect and limited payoff memory.}
    ({\bf Schematic illustrations}) On the left panels we show schematic
    illustrations of the perfect memory and the limited memory cases. The shaded
    background denotes the game phase information that an individual considered
    when updating strategies. In the case of perfect memory the entire region is
    shaded and in the case of limited memory only one turn with a single member
    of the population. ({\bf Simulations}) We have run four simulations of the
    pairwise comparison process for $T\!=\!10^7$ time steps. For each time step,
    we have recorded the current resident population ($y,p,q$). Since
    simulations are run for a relatively high continuation probability of
    $\delta\!=\!0.999$, we do not report the players' initial cooperation
    probability $y$. The graphs show how often the resident population chooses
    each combination ($p,q$) of conditional cooperation probabilities in the
    subsequent rounds. We also report the evolved cooperation rate which is
    calculated as the average cooperation rate within the resident population.
    ({\bf Perfect Memory}) In the case of low benefit the resident population
    either consists of defectors (with $p\!\approx\!q\!\approx\!0$) or of
    conditional cooperators. Conditional cooperators, or otherwise known as
    generous tit for tat, are a set of strategies that always cooperate
    following a cooperation ($p\!\approx\!1\!$) and cooperate with a probability
    $q$ given that the co-player has defected. $q$ denotes the generosity of a
    player. The resident population applies a conditional cooperator strategy
    for which $q\!\le\!1\!-\!c/b\!=\!0.67$. This is true also for the case of
    high benefit. In this case the population mainly consists of conditional
    cooperators of the form ($p\!\approx\!1\!, q\!\le\!1\!-\!1/10\!=\!0.9$). In
    the Supplementary Information Section 2.1 we show that a conditional
    cooperator needs to be of the form ($p\!\approx\!1\!, q\!\le\!1\!-\!c/b$) to
    not be invaded by defecting strategies. A higher generosity in the
    population results in a higher average cooperation rate. The average
    cooperation rate increases from 52\% for \(b\) to 98\% for \(b=10\). ({\bf
    Limited Memory}) When players update their strategies based on their
    realized payoffs in the last round, there are two different predominant
    behaviors regardless of the benefit value. The resident population either
    consists of defectors (with $p\!\approx\!q\!\approx\!0$) or of conditional
    cooperators. The maximum level of $q$ consistent with stable cooperation is
    somewhat smaller compared to the perfect memory setting, $q\!<\!0.5$.
    Namely, in the Supplementary Information Section 2.2.1 we show that
    regardless of the value of benefit, in the case of limited payoff memory, an
    individual needs a generosity smaller of $0.5$ to repel defectors. The plots
    are fairly similar regardless of the benefit, and the evolved cooperation
    rate only slightly increases from 37\% to 51\%. Parameters: $N\!=\!100$,
    $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.999$.}
    \label{fig:expected_and_stochastic_for_donation}
\end{figure}

We further explore the effects of the cooperation benefit and the strength of
selection on the evolved generosity \(q\) and cooperation rate
(Figure~\ref{fig:cooperation_rate_over_benefit_and_beta}).
Figure~\ref{fig:cooperation_rate_over_benefit_and_beta}\textbf{A} suggests that
perfect memory always yields a higher cooperation rate. We observe that the
cooperation rate increases as the value of the benefit gets higher, whereas in
comparison for the limited memory payoffs, the cooperation rate remains
unchanged (as we would have expected) at approximately 50\% once \(b=5\). From
Figure~\ref{fig:cooperation_rate_over_benefit_and_beta}\textbf{B} we observe
that for weak selection, \(\beta < 1\), the two methods yield similar results,
however, as \(\beta\) increases there is variation in the evolving populations.
In the case of expected payoffs the resident populations become more cooperative
as \(\beta\) increases, whereas in the case of limited memory payoffs, the
resident populations become more defective.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=.75\textwidth]{static/cooperation_rate_over_b_and_beta.pdf}
    \caption{{\bf The evolution of cooperation and generosity for different
    values of benefit (A) and strength of selection (B).} We report the average
    cooperation and the average reciprocity. The average cooperation rate is the
    average cooperation rate within the resident population. For the average
    reciprocity we select the residents that have a $p \approx 1$ and we take
    the average of their cooperation probability $q$. ({\bf A}) We vary the
    benefit of cooperation $b$. In all cases, perfect memory updating payoffs
    appear to overestimate the average cooperation rate the population achieves.
    As expected in the case of limited memory the average generosity over the
    different values of benefit remains the same ($q \approx 0.5$), and as a
    result so does the average cooperation. ({\bf B}) We vary the selection
    strength $\beta$. For weak selection, \(\beta < 1\), the two methods yield
    similar results. However, as \(\beta\) increases in the case of limited
    memory payoffs the resident populations become more defective. Unless
    explicitly varied, the parameters of the simulation are $N\!=\!100$,
    $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run
    for $T\!=\!5\times 10^7$ time steps for each parameter
    combination.}\label{fig:cooperation_rate_over_benefit_and_beta}
\end{figure}

@@

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\textwidth]{static/more_memory_summary_results.pdf}
  \caption{{\bf Average cooperation rates for different updating payoffs.} (From
  right to left) updating Payoffs based  on perfect memory, limited memory,
  the last round payoff of two interactions,
  the last two rounds payoff of one interaction and 
  the last two round payoffs of two interactions.({\bf High Benefit})
  ({\bf Low Benefit}) We vary the selection strength $\beta$. In all
  cases, stochastic payoff evaluation tends to reduce the evolving cooperation
  rates. Unless explicitly varied, the parameters of the simulation are
  $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$.
  Simulations are run for $T\!=\!5\times 10^7$ time steps for each parameter
  combination.}\label{fig:cooperation_rate_updating_payoffs}
\end{figure}
